Progress Report on the ``Adult Database'' analysis

### 2009-11-12 report/paper.tex
added cluster analysis.

\documentclass[a4paper]{llncs}
\usepackage{graphicx}
\usepackage{url}
\usepackage{longtable}
\usepackage[latin1]{inputenc}
\usepackage{subfigure}
\usepackage{a4wide}
\urldef{\mails}\path|{ei05011,ei05028}@fe.up.pt|
\bibliographystyle{splncs}
\begin{document}
\title{Progress Report on the ``Adult Database'' analysis}
\author{Flávio Cruz \and João Azevedo}
\institute{Faculdade de Engenharia da Universidade do Porto\\
    Rua Dr. Roberto Frias, s/n 4200-465 Porto PORTUGAL
    \mails}
\maketitle

@@

Another method to analyze the dataset is based on the generation of clusters. 
Clusters are sets of data in which objects in the same cluster are similar, and
objects in different clusters are dissimilar. The first attempt at clustering
used only three variables: age, education\_num and hours\_per\_week, the same
variables we saw that had a greater impact on the amount gained by the
individual. The data was normalized using the Z-Transformation method and the 
K-Means method was used to obtain the final clusters, with a k of 6. Figures 
\ref{fig:cluster_1_age_education} and \ref{fig:cluster_1_education_hpw} show
scatter plots of the variables, colored by clusters. One thing one is able to
note is that people with younger age belong to cluster \#2. In terms of
education, people with higher education belong to cluster \#0, while people with
lower education belong to cluster \#1. In terms of hours of work per week, 
distribution is rather similar, but cluster \#2 and \#3 have a greater amount of
lower and higher values, respectively. Figure \ref{fig:cluster_1_cluster_plus50}
shows how clusters relate to the class values. Although the distribution is 
balanced, clusters \#1 and \#2 have a greater amount of people gaining below 
50K. Notice, as stated before, that this were the cluster having people with
lower education and younger age, respective.

@@

@@

>>>>

\documentclass[a4paper]{llncs}
\usepackage{graphicx}
\usepackage{url}
\usepackage{longtable}
\usepackage[latin1]{inputenc}
\usepackage{subfigure}
\usepackage{a4wide}
\urldef{\mails}\path|{ei05011,ei05028}@fe.up.pt|
\bibliographystyle{splncs}
\begin{document}
\title{Progress Report on the ``Adult Database'' analysis}
\author{Flávio Cruz \and João Azevedo}
\institute{Faculdade de Engenharia da Universidade do Porto\\
    Rua Dr. Roberto Frias, s/n 4200-465 Porto PORTUGAL\\
    \mails}
\maketitle

@@

Another method to analyze the dataset is based on the generation of clusters. 
Clusters are sets of data in which objects in the same cluster are similar, and
objects in different clusters are dissimilar. The first attempt at clustering
used only three variables: age, education\_num and hours\_per\_week, the same
variables we saw that had a greater impact on the amount gained by the
individual. The data was normalized using the Z-Transformation method and the 
K-Means method was used to obtain the final clusters, with a k of 6. Figures 
\ref{fig:cluster_1_age_education}, \ref{fig:cluster_1_education_hpw} and \ref{fig:cluster_1_age_hpw} show
scatter plots of the variables, colored by clusters. One thing one is able to
note is that people with younger age belong to cluster \#2. In terms of
education, people with higher education belong to cluster \#0, while people with
lower education belong to cluster \#1. In terms of hours of work per week, 
distribution is rather similar, but cluster \#2 and \#3 have a greater amount of
lower and higher values, respectively. Figure \ref{fig:cluster_1_cluster_plus50}
shows how clusters relate to the class values. Although the distribution is 
balanced, clusters \#1 and \#2 have a greater amount of people gaining below 
50000 dollars. Notice, as stated before, that this were the clusters having 
people with lower education and younger age, respectively.

Clustering with the remaining attributes leads to results more difficult to
analyze. The same process was applied, using the Z-Transformation method for
normalization and the K-Means algorithm with a k of 6, but this time with all
the attributes. The distribution of values is more balanced, although a very
high value of capital gain, very unusual on the data set, leads to the creation
of a very well identified cluster, as can be seen on table \ref{tab:centroids}.
Particular values of race, workclass and native country also identify clusters
\#2, \#3 and \#2, respectively. Notice that this attributes were the ones
refered to before as having a given value more frequent than others, thus
leading to this particular composition of clusters.

@@

\begin{table}
\centering
\begin{tabular}{ | l | l | l | l | l | l | l | }
\hline
\textbf{Attribute} & \textbf{Centroid \#0} & \textbf{Centroid \#1} & \textbf{Centroid \#2} & \textbf{Centroid \#3} & \textbf{Centroid \#4} & \textbf{Centroid \#5} \\
\hline
age	& -0.874 & 0.692 & -0.166 & 0.246 & 0.570 & 0.190 \\
\hline
workclass & -0.272 & -0.063 & -0.036 & 2.208 & 0.455 & -0.364 \\
\hline
education\_num & -0.101 & -0.172 & -0.177 & 0.0309 & 1.103 & 0.102 \\
\hline
marital\_status & -0.679 & 1.693 & -0.083 & -0.150 & -0.007 & -0.108 \\
\hline
occupation & -0.305 & -0.303 & 0.091 & 0.952 & -0.434 & 0.010 \\
\hline
relationship & 0.703 & 0.526 & 0.189 & -0.053 & -0.351 & -0.507 \\
\hline
race & -0.086 & -0.020 & 2.630 & -0.147 & -0.002 & -0.230 \\
\hline
sex & 0.662 & 1.199 & 0.019 & -0.161 & -0.409 & -0.635 \\
\hline
capital\_gain & -0.116 & -0.091 & -0.086 & -0.029 & 13.394 & -0.040 \\
\hline
capital\_loss & -0.115 & -0.079 & -0.021 & 0.025 & -0.217 & 0.079 \\
\hline
hours\_per\_week & -0.523 & -0.231 & -0.031 &-0.071 & 0.758 & 0.339 \\
\hline
native\_country & -0.186 & -0.154 & 2.951 & -0.181 & -0.075 & -0.171 \\
\hline
\end{tabular}
\caption{Centroid table for the results of the application of the K-Means algorithm with a k of 6}
\label{tab:centroids}
\end{table}

@@

\begin{figure}
\centering
\includegraphics[width=10cm]{cluster_1_age_hpw.png}
\caption{Scatter plot of age and hours of work per week}
\label{fig:cluster_1_age_hpw}
\end{figure}
